---
title: "kaggle-titanic"
author: "Jason Murray"
date: '2016-12-10'
output: html_document
---

##Summary  
I'm new to kaggle and data anlytics in general.  This document covers my exploration of the titanic data for the kaggle challenge.  The goal is to build a model that will accurately predict survival for the test set of passengers based on learning from the training set.  
<br>

###Libraries
```{r libraries, message=FALSE}
library(mice)
library(dplyr)
library(ggplot2)
library(caret)
library(caretEnsemble)
library(parallel)
library(doParallel)
```

###Data download and load
```{r downLoad}
## Data was obtained from the kagggle challenge website
## https://www.kaggle.com/c/titanic/

##Train Data
trainURL <- "https://kaggle2.blob.core.windows.net/competitions-data/kaggle/3136/train.csv?sv=2015-12-11&sr=b&sig=sR2hRdWWjjygbvaWj3dCbjqBPzjrJ4gAQulu3w1xw2M%3D&se=2016-12-14T00%3A56%3A01Z&sp=r"
trainfile <- "./data/train.csv"

##Test Data
testURL <- "https://kaggle2.blob.core.windows.net/competitions-data/kaggle/3136/test.csv?sv=2015-12-11&sr=b&sig=cD%2BWsR6UgfAdMWvjquCBnQCYzVgJqDvL85fXpngwUko%3D&se=2016-12-14T00%3A56%3A33Z&sp=r"
testfile <- "./data/test.csv"

##Download 
#download.file(trainURL,trainfile)
#download.file(testURL,testfile)

##Read
trainData <- read.csv(trainfile, na.strings = c("", " "), stringsAsFactors = FALSE)
testData <- read.csv(testfile, na.strings = c("", " "), stringsAsFactors = FALSE)

##Combine test and train data for initial exploration and imputation.
allData <- bind_rows(trainData, testData)
```
<br>

##Data Cleaning
<br>

For reference here is the information about the data provided by the challenge
<br>

|Variable | Description|
|---------|------------|
|survival | Survival   
|	  | (0 = No; 1 = Yes) 
|pclass   | Passenger Class 
|         | (1 = 1st; 2 = 2nd; 3 = 3rd)
|name     | Name
|sex      | Sex
|age      | Age
|sibsp    | Number of Siblings/Spouses Aboard
|parch    | Number of Parents/Children Aboard
|ticket   | Ticket Number
|fare     | Passenger Fare
|cabin    | Cabin
|embarked | Port of Embarkation
|         | (C = Cherbourg; Q = Queenstown; S = Southampton)

SPECIAL NOTES:
Pclass is a proxy for socio-economic status (SES)
 1st ~ Upper; 2nd ~ Middle; 3rd ~ Lower

Age is in Years; Fractional if Age less than One (1)
 If the Age is Estimated, it is in the form xx.5

With respect to the family relation variables (i.e. sibsp and parch)
some relations were ignored.  The following are the definitions used
for sibsp and parch.

Sibling:  Brother, Sister, Stepbrother, or Stepsister of Passenger Aboard Titanic
Spouse:   Husband or Wife of Passenger Aboard Titanic (Mistresses and Fiances Ignored)
Parent:   Mother or Father of Passenger Aboard Titanic
Child:    Son, Daughter, Stepson, or Stepdaughter of Passenger Aboard Titanic

Other family relatives excluded from this study include cousins,
nephews/nieces, aunts/uncles, and in-laws.  Some children travelled
only with a nanny, therefore parch=0 for them.  As well, some
travelled with very close friends or neighbors in a village, however,
the definitions do not support such relations.

***


#####Then we take a first look
```{r}
str(allData)
```
<br>

Might be helpful to split out the Last Name and Title from the name column.  Let's try it.

```{r}
allData <- allData %>% 
	mutate(lastname = gsub('(.*),.*','\\1',Name)) %>%
	mutate(title = gsub('.*, (\\w+).\\s.*','\\1',Name))

table(allData$title)
```

So there is likely not much information in the title.  It mostly shows male/female distinction and age which we already have.  There are a few unique titles but there aren't enough in each category to be valuable.  Last name could provide some information as it could link people together in families.  

Let's look for NA values
```{r}
sapply(allData, function(x) sum(is.na(x)))
```

So let's take a look at the missing fair.
```{r}
subset(allData, is.na(Fare))
```

Let's set the fare to to the average fair for Plcass of 3 and call it done.  
```{r}
p3Fares <- subset(allData, Pclass==3)
allData[1044,"Fare"] <- mean(p3Fares$Fare, na.rm = TRUE)
```


The ages will take something more with 263 missing values.  21% missing is a lot but let's try and impute the missing values.
```{r}
##Select the data we want to use for imputation. 
vars <- c("Pclass", "Sex", "SibSp", "Parch", "Fare", "Embarked", "Age")
allDataMice <- allData %>% select(one_of(vars)) 
mice_out <- mice(allDataMice)
mice_values <- complete(mice_out)

allData$Age <- mice_values$Age
```



Looking at the cabin info it's letter and numeric probably indicating deck and room number.  There are 1014 missing values so this is not going to be of value.

Ticket numbers are all over the place and I don't see it really having much value for prediction.  

Embarked has a few missing values
```{r}
sapply(allData, function(x) sum(is.na(x)))
```

Let's take a look at these and see what we can do.

```{r}
allData %>% filter(is.na(Embarked))
#allData <- allData %>% filter(!is.na(Embarked))
```

Let's see where other passengers in their class embarked.
```{r}
table(allData %>% filter(Pclass==1) %>% select(Embarked))

```
Let's call them both S since that's where the majority come from

```{r}
allData[62,"Embarked"] <- "S"
allData[830,"Embarked"] <- "S"
```



So after looking at all the variables I'm going to use Pclass, SibSp, Parch, title, Fare, Sex, Embarked as they all may have some information and are complete to build a random forest model and see how it does.   

#####Factorizing variables
```{r}
##Factorizing
allData$Survived <- as.factor(allData$Survived)
allData$Pclass <- as.factor(allData$Pclass)
allData$Sex <- as.factor(allData$Sex)
allData$title <- as.factor(allData$title)
allData$Embarked <- as.factor(allData$Embarked)
#allData$Parch <- as.factor(allData$Parch)
#allData$SibSp <- as.factor(allData$SibSp)

##Adding some valid factor labels
levels(allData$Pclass) <- c("First","Second","Third")
levels(allData$Survived) <- c("Perished","Survived")

```


####Seperating out factors

Now I'll seperate out the factors I've decided to use and then resplit the data into train and test sets.  

```{r}
factorsToUse <- c("PassengerId", "Survived", "Pclass", "Sex", "SibSp", "Parch", "Fare", "Embarked", "Age") 
	
allDataSF <- allData %>% select(one_of(factorsToUse))
titanicTest <-  allDataSF %>% filter(is.na(Survived))
titanicTrain <- allDataSF %>% filter(!is.na(Survived)) %>% select(-PassengerId)
```


##Building a model

####Parallel processing
```{r, message=FALSE, warning=FALSE}
## Set up parallel processing
cluster <- makeCluster(detectCores() - 1) 
registerDoParallel(cluster)
```

####Build model
```{r, cache=TRUE, message=FALSE, warning=FALSE}

set.seed(42)

rf_ctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 5, allowParallel=TRUE)
rf_tunegrid <- expand.grid(.mtry=c(1:6))

model_rf <- train(Survived ~ . , data=titanicTrain, method='rf', preProcess = c("center", "scale"), trControl = rf_ctrl, tuneGrid = rf_tunegrid)
```

####Make Predictions


```{r, cache=TRUE, message=FALSE, warning=FALSE}
titanicPredictions_rf <- predict(model_rf,newdata=titanicTest)
titanicSubmission_rf <- data_frame(PassengerID = titanicTest$PassengerId, Survived = as.numeric(titanicPredictions_rf)-1)
write.csv(titanicSubmission_rf, file = "titanicSubmission_rf.csv", row.names = FALSE)
```

####Build another model
```{r, cache=TRUE, message=FALSE, warning=FALSE}

set.seed(42)

svm_ctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 5, allowParallel=TRUE)
model_svm <- train(Survived ~ . , data=titanicTrain, method='svmRadial', preProcess = c("center", "scale"), trControl = svm_ctrl, tuneLength = 15 )
```

####Make Predictions with it.

```{r, cache=TRUE, message=FALSE, warning=FALSE}
titanicPredictions_svm <- predict(model_svm,newdata=titanicTest)
titanicSubmission_svm <- data_frame(PassengerID = titanicTest$PassengerId, Survived = as.numeric(titanicPredictions_svm)-1)
write.csv(titanicSubmission, file = "titanicSubmission_svm.csv", row.names = FALSE)
```

####Build another model
```{r, cache=TRUE, message=FALSE, warning=FALSE}

set.seed(42)

xgbLinear_ctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 5, allowParallel=TRUE)
model_xgbLinear <- train(Survived ~ . , data=titanicTrain, method='xgbLinear', preProcess = c("center", "scale"), trControl = xgbLinear_ctrl)
```
####Make Predictions with it.

```{r, cache=TRUE, message=FALSE, warning=FALSE}
titanicPredictions_xgblinear <- predict(model_xgbLinear,newdata=titanicTest)
titanicSubmission_xgblinear <- data_frame(PassengerID = titanicTest$PassengerId, Survived = as.numeric(titanicPredictions_xgblinear)-1)
write.csv(titanicSubmission, file = "titanicSubmission_xgblinear.csv", row.names = FALSE)
```


####Try combining the models
```{r, cache=TRUE, message=FALSE, warning=FALSE}
set.seed(42)
##Tuning
stacked_ctrl <- trainControl(method = "repeatedcv", number = 10, repeats = 5, allowParallel=TRUE, classProbs=TRUE)

##Build Models
models <- caretList(Survived ~ . , data=titanicTrain, trControl = stacked_ctrl,
			   tuneList = list(
			   	rf=caretModelSpec(method='rf', tuneGrid=rf_tunegrid),
			   	svmRadial=caretModelSpec(method='svmRadial', tuneLength=15),
			   	xgbLinear=caretModelSpec(method='xgbLinear')
			   ))
##Combine them
stacked_tunegrid <- expand.grid(.mtry=c(1:6))
models_stacked <- caretStack(models, method="rf", metric="Accuracy", trControl=stacked_ctrl, tuneGrid=stacked_tunegrid)

```


####Make predictions with the stacked model
```{r, cache=TRUE, message=FALSE, warning=FALSE}
titanicPredictions_stacked <- predict(models_stacked,newdata=titanicTest)
titanicSubmission_stacked <- data_frame(PassengerID = titanicTest$PassengerId, Survived = as.numeric(titanicPredictions_stacked)-1)
write.csv(titanicSubmission, file = "titanicSubmission_stacked.csv", row.names = FALSE)
```

```{r}

```

